---
title: "Quick start with abcrlda"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Abcrlda}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Introduction 

The abcrlda package implements Asymptotically Bias-Corrected Regularized Linear Discriminant Analysis.
It is an extension of classical RLDA (Regularized Linear Discriminant Analysis) algorithm 
that improves performance in case of Cost-Sensitive Binary Classification.
ABCRLDA makes it possible to prioritize classes differently during fitting process.
Missclassification cost could be different for different classes.


For some application cost of making mistake
For example, in case of some diseases it is more important to
decrease number of false negatives (illness is not detected) even at expence 


## Quick Start
The purpose of this section is to give examples of basic usage in order to give users a general sense of the package.
For illustration purposes iris dataset will be used.
First, we load the abcrlda package and iris dataset

```{r}
library(abcrlda)
data(iris)
```

abcrlda is designed for binary classification therefore we will limit training data to two species (virginica and versicolor) from iris dataset

```{r}
train_data <- iris[which(iris[, ncol(iris)] == "versicolor" |
                         iris[, ncol(iris)] == "virginica"), 1:4]
train_label <- factor(iris[which(iris[, ncol(iris)] == "versicolor" |
                                 iris[, ncol(iris)] == "virginica"), 5])
```

```{r}
model <- abcrlda(train_data, train_label)
predict(model, train_data)
```
### Controlling priorities
```{r}
model <- abcrlda(train_data, train_label, cost = c(0.75, 0.25))


test_versicolor <- iris[which(iris[, ncol(iris)] == "versicolor"), 1:4]
test_virginica <- iris[which(iris[, ncol(iris)] == "virginica"), 1:4]

r <- predict(model, test_versicolor)
sum(r != "versicolor") / nrow(test_virginica)  # resubstitution error for first class

r <- predict(model, test_virginica)
sum(r != "virginica") / nrow(test_virginica)  # resubstitution error for second class

```

If costs are switched


```{r}
model <- abcrlda(train_data, train_label, cost = c(0.25, 0.75))


test_versicolor <- iris[which(iris[, ncol(iris)] == "versicolor"), 1:4]
test_virginica <- iris[which(iris[, ncol(iris)] == "virginica"), 1:4]

r <- predict(model, test_versicolor)
sum(r != "versicolor") / nrow(test_virginica)  # resubstitution error for first class

r <- predict(model, test_virginica)
sum(r != "virginica") / nrow(test_virginica)  # resubstitution error for second class
```


### Affects of cost values on classification

```{r}
model <- abcrlda(train_data, train_label, gamma = 0.5, cost = 0.75)
a <- predict(model, train_data)
# same params but more explicit
model <- abcrlda(train_data, train_label, gamma = 0.5, cost = c(0.75, 0.25))
b <- predict(model, train_data)
# same class costs ratio
model <- abcrlda(train_data, train_label, gamma = 0.5, cost = c(3, 1))
c <- predict(model, train_data)
# all this model will give the same predictions
all(a == b & a == c & b == c)
```

### Grid Search for optimal Hyperparameters
```{r}
cost_range <- seq(0.1, 0.9, by = 0.2)
gamma_range <- c(0.1, 1, 10, 100, 1000)

gs <- grid_search(train_data, train_label,
                  range_gamma = gamma_range,
                  range_cost = cost_range,
                  method = "estimator")
model <- abcrlda(train_data, train_label,
                 gamma = gs$gamma, cost = gs$cost)
# predict(model, train_data)
gs
```

```{r eval=T, include=T}
cost_range <- matrix(1:8, ncol = 2)
gamma_range <- c(1, 10)

gs <- grid_search(train_data, train_label,
                  range_gamma = gamma_range,
                  range_cost = cost_range,
                  method = "cross")
model <- abcrlda(train_data, train_label,
                 gamma = gs$gamma, cost = gs$cost)
# predict(model, train_data)
gs
```
