---
title: "Quick start with abcrlda"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Abcrlda}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Introduction 

The abcrlda package implements Asymptotically Bias-Corrected Regularized Linear Discriminant Analysis.
It is an extension of classical RLDA (Regularized Linear Discriminant Analysis) algorithm 
that improves performance in case of Cost-Sensitive Binary Classification.
ABCRLDA makes it possible to prioritize classes differently during fitting process.
Missclassification cost could be different for different classes.


```{r eval=F, include=F}
For some application cost of making mistake
For example, in case of some diseases it is more important to
decrease number of false negatives (illness is not detected) even at expence 
```

## Quick Start
The purpose of this section is to give examples of basic usage in order to give users a general sense of the package.
For illustration purposes iris dataset will be used.
First, we load the abcrlda package and iris dataset

```{r}
library(abcrlda)
data(iris)
```

Abcrlda is designed for binary classification therefore we will limit training data to two species (virginica and versicolor), 50 observation for each

```{r}
train_data <- iris[which(iris[, ncol(iris)] == "versicolor" |
                         iris[, ncol(iris)] == "virginica"), 1:4]
train_label <- factor(iris[which(iris[, ncol(iris)] == "versicolor" |
                                 iris[, ncol(iris)] == "virginica"), 5])
```

After that we are ready to fit model and make predictions using resubstitution
```{r}
model <- abcrlda(train_data, train_label)
predict(model, train_data)
```

Accuracy can be evaluated by this function
```{r}
risk_calculate(model, train_data, train_label)
```
### Controlling priorities

Cost is one of the most important hyperparameters for the abcrlda. 
It controls prioritization of classes during model fitting and it used for overall risk calculation.
Error rate for more important class tends to be smaller. 

`R = e_0 * cost_10 + e_1 * cost_01`




```{r}
model <- abcrlda(train_data, train_label, cost = c(0.75, 0.25))


test_versicolor <- iris[which(iris[, ncol(iris)] == "versicolor"), 1:4]
test_virginica <- iris[which(iris[, ncol(iris)] == "virginica"), 1:4]

r <- predict(model, test_versicolor)
sum(r != "versicolor") / nrow(test_virginica)  # resubstitution error for first class

r <- predict(model, test_virginica)
sum(r != "virginica") / nrow(test_virginica)  # resubstitution error for second class

```

If costs are reversed error rates are changed accordingly:

```{r}
model <- abcrlda(train_data, train_label, cost = c(0.25, 0.75))


test_versicolor <- iris[which(iris[, ncol(iris)] == "versicolor"), 1:4]
test_virginica <- iris[which(iris[, ncol(iris)] == "virginica"), 1:4]

r <- predict(model, test_versicolor)
sum(r != "versicolor") / nrow(test_virginica)  # resubstitution error for first class

r <- predict(model, test_virginica)
sum(r != "virginica") / nrow(test_virginica)  # resubstitution error for second class
```


### Affects of cost values on classification

Cost values can be provided in different ways.
There is short form for normalized cost values:

```{r}
model <- abcrlda(train_data, train_label, gamma = 0.5, cost = 0.75)
a <- predict(model, train_data)
```

Which is equivalent to:

```{r}
# same params but more explicit
model <- abcrlda(train_data, train_label, gamma = 0.5, cost = c(0.75, 0.25))
b <- predict(model, train_data)
```

Scaling of a cost do not change prediction that will be made by model: 

```{r}
# same class costs ratio
model <- abcrlda(train_data, train_label, gamma = 0.5, cost = c(3, 1))
c <- predict(model, train_data)

# all this model will give the same predictions
all(a == b & a == c & b == c)
```

### Grid Search for optimal Hyperparameters

This package contain function to performs grid search for optimal hyperparameters 
within specified parameters range. 
Performanse of a model can be evaluated using double asymptotic risk estimation or cross validation for separate sampling adjasted for cost.
Double asymptotic risk estimation is faster option because it uses closed form formula for risk estimation
while cross validation is more computationally expensive.

```{r eval=F, include=T}
cost_range <- seq(0.1, 0.9, by = 0.2)
gamma_range <- c(0.1, 1, 10, 100, 1000)

gs <- grid_search(train_data, train_label,
                  range_gamma = gamma_range,
                  range_cost = cost_range,
                  method = "estimator")
model <- abcrlda(train_data, train_label,
                 gamma = gs$gamma, cost = gs$cost)

```

```{r eval=F, include=T}
cost_range <- matrix(1:10, ncol = 2)
gamma_range <- c(0.1, 1, 10, 100, 1000)

gs <- grid_search(train_data, train_label,
                  range_gamma = gamma_range,
                  range_cost = cost_range,
                  method = "cross")
model <- abcrlda(train_data, train_label,
                 gamma = gs$gamma, cost = gs$cost)
```
